Biological shape as a weighted network

### 2013-01-04 documents/report_1.tex
initial commit. restarted analysis to be more elegant

\documentclass{article}
\frenchspacing

<<set-options, echo=FALSE, cache=FALSE, message=FALSE>>=
require(knitr)
require(formatR)
options(keep.comment = FALSE)
read_chunk('initial_outline.R')
read_chunk('landmark_montecarlo.R')
@

\SweaveOpts{external = TRUE, size='scriptsize'}

@@

I've implemented this Monte Carlo process as a few functions in R.
The core function is {\tt{jitter.landmark()}}.
<<jitter>>=
@

This functions are called {\tt{nsim}} times (deafult 1000) by the {\tt{riem.distribution()}} function.
Using the original data set, {\tt{jitter.landmark()}} is called on {\tt{nsim}} two shape combinations.
A lambda function is used to calculate the reimanian distance for all of the {\tt{nsim}} combinations.
Currently, the lambda function executed as part of a call to {\tt{mclapply()}} which is a parallel function from the {\tt{parallel}} core package.
Currently, it will use the maximum number of cores avaliable and the seed is optimzed for reproducibility.
Without this feature, the process will run much slower.
This function will not work on non-POSIX systems (i.e. Windows).
<<riem-dist>>=
@

@@

The interface function {\tt{shape.simulation()}} outputs a quantile value necessary for determining our ``at random.'' The deafult quantile of 0.50 is the same as the median of a sample.
<<shape-sim>>=
@

@@

\subsection{Fish example}
The first step is to import the correct libraries and data files.
The functions described above are part of my script file, while the {\tt{readtps.R}} file is my fork of Ananda Mahto's original script GIST LINK.
This fork improves integration of .tps files and the {\tt{shapes}} package.
Additionally, becreporause this is a randomization process, for reproducibiliity it is necessary to manually set the seed value. Additionally, the RNG kind is changed to the L'Ecuyer algorithm CITATION which is the best for random number generationalgorithm implemented in R for parallel processes.
<<fish-set-up, message=FALSE>>=
@

Next, we use GPA to fit our entire sample of \Sexpr{dim(fish)[3]} fish. We also calculate the pairwise \(\rho\) matrix. This is actually really dirty code that can probably be implemented as a single {\tt{outer}} call, but I've yet to figure it out.

<<fish-fit, eval=TRUE, cache=TRUE>>=
@

@@

<<fish-ex1, echo=FALSE, eval=FALSE, message=FALSE, fig.height= 4, results='hide'>>=
require(ggplot2)
# turn fish.fit$rotated into a two column matrix
fish.mean <- as.data.frame(fish.fit$mshape)
fish.fix <- fish.fit$rotated[, , -42]
fish.gg <- fish.fit$rotated[, , 1]
for(ii in 2:(dim(fish.fix)[3])) {
  fish.gg <- rbind(fish.gg, fish.fix[, , ii])
}

fish.gg <- as.data.frame(fish.gg)
fs <- ggplot(fish.gg, aes(x= V1, y= V2)) + geom_point(size = 1)
fs <- fs + scale_x_continuous(name = "X")
fs <- fs + scale_y_continuous(name = "Y")
fs <- fs + opts(panel.background = theme_rect(fill = "transparent",colour = NA), 
                panel.grid.minor = theme_blank(), 
                panel.grid.major = theme_blank(),
                axis.line = theme_segment(),
                axis.text.x = theme_text(colour = "black", size = 15),
                axis.text.y = theme_text(colour = "black", size = 15)) 
fs
@

<<fish-ex2, echo=FALSE, eval=FALSE, message=FALSE, fig.height= 4, results='hide'>>=
fish.eigsc <- as.data.frame(fish.fit$stdscores)
fam <- c(rep("Ano", times= length(ano$ID)), rep("Cur", times= length(cur$ID)))
fish.eigsc$fam <- fam
fish.eigsc <- fish.eigsc[-42, ]
fe <- ggplot(fish.eigsc, aes(x= PC1, y= PC2, colour= fam))
fe <- fe + geom_point(size = 3)
fe <- fe + scale_x_continuous(name = paste("PC1; ", 
                                           format(fish.fit$percent[1], digits= 3),
                                           sep= ''))
fe <- fe + scale_y_continuous(name = paste("PC2; ", 
                                           format(fish.fit$percent[2], digits= 3),
                                           sep= ''))
fe <- fe + opts(panel.background = theme_rect(fill = "transparent",colour = NA), 
                panel.grid.minor = theme_blank(), 
                panel.grid.major = theme_blank(),
                  axis.line = theme_segment(),
                axis.text.x = theme_text(colour = "black", size = 15),
                axis.text.y = theme_text(colour = "black", size = 15))
fe
@

<<fish-ex3, echo=FALSE, eval=FALSE, message=FALSE, fig.height= 4, results='hide'>>=
shapepca(fish.fit, joinline=c(1:35, 1))
@

<<fish-ex4, echo=FALSE, eval=FALSE, message=FALSE, fig.height= 4, results='hide'>>=
par(mfrow = c(1, 2))
tpsgrid(fish.fit$rotated[, , 1], fish.fit$mshape)
tpsgrid(fish.fit$rotated[, , 19], fish.fit$mshape)
@

\begin{figure}
\includegraphics[height=0.3\textheight]{figure/fish-ex1}
\caption{Landmark positions of \Sexpr{dim(fish.fit$rotated)[3] - 1} of the fish taxa. One specimen is excluded.\label{fish-plotshapes}}
\end{figure}

@@

<<fish-MC, eval=FALSE>>=
@

@@

<<fish-network, eval=FALSE>>=
@

@@

<<fish-plots, eval=FALSE, echo=FALSE>>=
@

@@

\newpage
For reproducibility information.
<<reproducibility>>=
sessionInfo()
@

>>>>

\documentclass{article}\usepackage{graphicx, color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\definecolor{fgcolor}{rgb}{0.2, 0.2, 0.2}
\newcommand{\hlnumber}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlfunctioncall}[1]{\textcolor[rgb]{0.501960784313725,0,0.329411764705882}{\textbf{#1}}}%
\newcommand{\hlstring}[1]{\textcolor[rgb]{0.6,0.6,1}{#1}}%
\newcommand{\hlkeyword}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlargument}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlcomment}[1]{\textcolor[rgb]{0.180392156862745,0.6,0.341176470588235}{#1}}%
\newcommand{\hlroxygencomment}[1]{\textcolor[rgb]{0.43921568627451,0.47843137254902,0.701960784313725}{#1}}%
\newcommand{\hlformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hleqformalargs}[1]{\textcolor[rgb]{0.690196078431373,0.250980392156863,0.0196078431372549}{#1}}%
\newcommand{\hlassignement}[1]{\textcolor[rgb]{0,0,0}{\textbf{#1}}}%
\newcommand{\hlpackage}[1]{\textcolor[rgb]{0.588235294117647,0.709803921568627,0.145098039215686}{#1}}%
\newcommand{\hlslot}[1]{\textit{#1}}%
\newcommand{\hlsymbol}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlprompt}[1]{\textcolor[rgb]{0.2,0.2,0.2}{#1}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\frenchspacing

@@

I've implemented this Monte Carlo process as a few functions in R.
The core function is {\tt{jitter.landmark()}}.
\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
jitter.landmark <- \hlfunctioncall{function}(land) \{
    \hlfunctioncall{for} (ii in \hlfunctioncall{seq}(\hlfunctioncall{ncol}(land))) \{
        land[, ii] <- land[, ii] + \hlfunctioncall{runif}(n = \hlfunctioncall{length}(land[, ii]), min = -(\hlfunctioncall{min}(\hlfunctioncall{abs}(\hlfunctioncall{diff}(land[, 
            ii])))), max = \hlfunctioncall{min}(\hlfunctioncall{abs}(\hlfunctioncall{diff}(land[, ii]))))
    \}
    land
\}
\end{alltt}
\end{kframe}
\end{knitrout}

This functions are called {\tt{nsim}} times (deafult 1000) by the {\tt{riem.distribution()}} function.
Using the original data set, {\tt{jitter.landmark()}} is called on {\tt{nsim}} two shape combinations.
A lambda function is used to calculate the reimanian distance for all of the {\tt{nsim}} combinations.
Currently, the lambda function executed as part of a call to {\tt{mclapply()}} which is a parallel function from the {\tt{parallel}} core package.
Currently, it will use the maximum number of cores avaliable and the seed is optimzed for reproducibility.
Without this feature, the process will run much slower.
This function will not work on non-POSIX systems (i.e. Windows).
\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
riem.distribution <- \hlfunctioncall{function}(fit.land, nsim = 1000) \{
    ran.land <- \hlfunctioncall{vector}(mode = \hlstring{"list"}, length = nsim)
    \hlfunctioncall{for} (ii in \hlfunctioncall{seq}(nsim)) \{
        s <- \hlfunctioncall{sample}(fit.land$n, 2)
        sh1 <- \hlfunctioncall{jitter.landmark}(fit.land$rotated[, , s[1]])
        sh2 <- \hlfunctioncall{jitter.landmark}(fit.land$rotated[, , s[2]])
        ran.land[[ii]] <- \hlfunctioncall{abind}(sh1, sh2, along = 3)
    \}
    res <- \hlfunctioncall{mclapply}(ran.land, \hlfunctioncall{function}(x) \{
        \hlfunctioncall{riemdis}(x[, , 1], x[, , 2])
    \}, mc.cores = \hlfunctioncall{detectCores}(), mc.set.seed = T)
    res
\}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

The interface function {\tt{shape.simulation()}} outputs a quantile value necessary for determining our ``at random.'' The deafult quantile of 0.50 is the same as the median of a sample.
\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
shape.simulate <- \hlfunctioncall{function}(fit.land, nsim = 1000, probs = 0.5) \{
    ri.di <- \hlfunctioncall{riem.distribution}(fit.land, nsim = nsim)
    \hlfunctioncall{quantile}(\hlfunctioncall{unlist}(ri.di), probs = probs)
\}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\subsection{Fish example}
The first step is to import the correct libraries and data files.
The functions described above are part of my script file, while the {\tt{readtps.R}} file is my fork of Ananda Mahto's original script GIST LINK.
This fork improves integration of .tps files and the {\tt{shapes}} package.
Additionally, becreporause this is a randomization process, for reproducibiliity it is necessary to manually set the seed value. Additionally, the RNG kind is changed to the L'Ecuyer algorithm CITATION which is the best for random number generationalgorithm implemented in R for parallel processes.
\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{require}(shapes)
\hlfunctioncall{require}(igraph)
\hlfunctioncall{require}(parallel)
\hlfunctioncall{require}(abind)
\hlfunctioncall{source}(\hlstring{"landmark_montecarlo.R"})
\hlfunctioncall{source}(\hlstring{"/home/peter/Documents/projects/3294254/readtps.R"})
\hlfunctioncall{RNGkind}(\hlstring{"L'Ecuyer-CMRG"})
\hlfunctioncall{set.seed}(1)
ano <- \hlfunctioncall{read.tps}(\hlstring{"AnostomidaeFinal.TPS"}, shapes = TRUE)
cur <- \hlfunctioncall{read.tps}(\hlstring{"CurimatidaeF.TPS"}, shapes = TRUE)
fish <- \hlfunctioncall{abind}(ano$landmark, cur$landmark)
\end{alltt}
\end{kframe}
\end{knitrout}

Next, we use GPA to fit our entire sample of 67 fish. We also calculate the pairwise \(\rho\) matrix. This is actually really dirty code that can probably be implemented as a single {\tt{outer}} call, but I've yet to figure it out.

\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
fish.fit <- \hlfunctioncall{procGPA}(fish)
riem.comp <- \hlfunctioncall{matrix}(0, ncol = fish.fit$n, nrow = fish.fit$n)
samp <- fish.fit$n
\hlfunctioncall{for} (ii in \hlfunctioncall{seq}(samp)) \hlfunctioncall{for} (kk in \hlfunctioncall{seq}(samp)) \{
    riem.comp[ii, kk] <- \hlfunctioncall{riemdist}(fish.fit$rotated[, , ii], fish.fit$rotated[, 
        , kk])
\}
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{figure}
\includegraphics[height=0.3\textheight]{figure/fish-ex1}
\caption{Landmark positions of 66 of the fish taxa. One specimen is excluded.\label{fish-plotshapes}}
\end{figure}

@@

\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
cutoff <- \hlfunctioncall{shape.simulate}(fish.fit, nsim = 10, probs = 0.5)
riem.comp[riem.comp > cutoff] <- 0
riem.comp[riem.comp != 0] <- 1/riem.comp[riem.comp != 0]
\end{alltt}
\end{kframe}
\end{knitrout}

@@

\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
data.graph <- \hlfunctioncall{graph.adjacency}(riem.comp, mode = \hlstring{"undirected"}, weighted = TRUE)
data.fc <- \hlfunctioncall{fastgreedy.community}(data.graph)
data.ebc <- \hlfunctioncall{edge.betweenness.community}(data.graph)
data.lec <- \hlfunctioncall{leading.eigenvector.community}(data.graph)
\end{alltt}
\end{kframe}
\end{knitrout}

@@

@@

\newpage
For reproducibility information.
\begin{knitrout}\scriptsize
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlfunctioncall{sessionInfo}()
\end{alltt}
\begin{verbatim}
## R version 2.15.1 (2012-06-22)
## Platform: x86_64-pc-linux-gnu (64-bit)
## 
## locale:
##  [1] LC_CTYPE=en_US.UTF-8       LC_NUMERIC=C              
##  [3] LC_TIME=en_AU.UTF-8        LC_COLLATE=en_US.UTF-8    
##  [5] LC_MONETARY=en_AU.UTF-8    LC_MESSAGES=en_US.UTF-8   
##  [7] LC_PAPER=C                 LC_NAME=C                 
##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
## [11] LC_MEASUREMENT=en_AU.UTF-8 LC_IDENTIFICATION=C       
## 
## attached base packages:
## [1] parallel  stats     graphics  grDevices utils     datasets  methods  
## [8] base     
## 
## other attached packages:
## [1] abind_1.4-0          igraph_0.6-2         shapes_1.1-5        
## [4] MASS_7.3-20          rgl_0.92.892         scatterplot3d_0.3-33
## [7] formatR_0.6          knitr_0.7           
## 
## loaded via a namespace (and not attached):
## [1] digest_0.5.2   evaluate_0.4.2 plyr_1.7.1     stringr_0.6.1 
## [5] tools_2.15.1  
\end{verbatim}
\end{kframe}
\end{knitrout}
